---
title: "apan5200kaggle1"
author: "Yosue Canaviri"
date: "2023-03-09"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(doParallel)
library(foreach)
library(writexl)
library(stringr)
library(caTools)
library(ranger)
library(doParallel)
```



```{r}
dat <-read.csv("analysisData.csv") # Training set
test <-read.csv("scoringData.csv") # Testing Set
```


```{r}
#Creating track_minutes
dat<- dat%>%mutate(track_minutes=track_duration/(1000*60))


dat$time_sig_cat<- character(length= nrow (dat))
dat$time_sig_cat[dat$time_signature==4]<-"Four"
dat$time_sig_cat[dat$time_signature==3]<-"Three"
dat$time_sig_cat[dat$time_signature %in% c(0,1,5)]<-"Other"

dat$key_cat<- character(length= nrow (dat))
dat$key_cat[dat$key==0]<-"Zero"
dat$key_cat[dat$key==1]<-"One"
dat$key_cat[dat$key==2]<-"Two"
dat$key_cat[dat$key==3]<-"Three"
dat$key_cat[dat$key==4]<-"Four"
dat$key_cat[dat$key==5]<-"Five"
dat$key_cat[dat$key==6]<-"Six"
dat$key_cat[dat$key==7]<-"Seven"
dat$key_cat[dat$key==8]<-"Eight"
dat$key_cat[dat$key==9]<-"Nine"
dat$key_cat[dat$key==10]<-"Ten"
dat$key_cat[dat$key==11]<-"Eleven"

#Taking the numbers of words in each song
dat$song1 <- gsub("\\([^()]*\\)", "", dat$song)
dat$song_counter <- str_count(dat$song1, "\\s+") + 1
#Transforming the count to characters to make them categorical
dat$song_count[dat$song_counter==1]<-"One"
dat$song_count[dat$song_counter==2]<-"Two"
dat$song_count[dat$song_counter==3]<-"Three"
dat$song_count[dat$song_counter==4]<-"Four"
dat$song_count[dat$song_counter==5]<-"Five"
dat$song_count[dat$song_counter==6]<-"Six"
dat$song_count[dat$song_counter==7]<-"Seven"
dat$song_count[dat$song_counter >=8]<-"Eight Plus"

#Taking the numbers of words in each performer
dat$performer_counter <- str_count(dat$performer, "\\s+") + 1
#Transforming the count to characters to make them categorical
dat$performer_count[dat$performer_counter==1]<-"One"
dat$performer_count[dat$performer_counter==2]<-"Two"
dat$performer_count[dat$performer_counter==3]<-"Three"
dat$performer_count[dat$performer_counter>=4]<-"Four Plus"


#Categorizing the genres
dat$pop_genre <- grepl("pop", dat$genre, ignore.case = TRUE)
dat$rock_genre <- grepl("rock", dat$genre, ignore.case = TRUE)
dat$adult_genre <- grepl("adult standard", dat$genre, ignore.case = TRUE)
dat$soul_genre <- grepl("soul", dat$genre, ignore.case = TRUE)
dat$rap_genre <- grepl("rap", dat$genre, ignore.case = TRUE)
dat$country_genre <- grepl("country", dat$genre, ignore.case = TRUE)
dat$hiphop_genre <- grepl("hip hop", dat$genre, ignore.case = TRUE)
dat$electro_genre <- grepl("electro", dat$genre, ignore.case = TRUE)
dat$latin_genre <- grepl("latin", dat$genre, ignore.case = TRUE)

dat$mode<-as.factor(dat$mode)

#Converting booleans to categorical
dat <- dat %>% 
  mutate_if(is.logical, as.factor)
```

  
```{r}
#Creating track_minutes
test<- test%>%mutate(track_minutes=track_duration/(1000*60))


test$time_sig_cat<- character(length= nrow (test))
test$time_sig_cat[test$time_signature==4]<-"Four"
test$time_sig_cat[test$time_signature==3]<-"Three"
test$time_sig_cat[test$time_signature %in% c(0,1,5)]<-"Other"

test$key_cat<- character(length= nrow (test))
test$key_cat[test$key==0]<-"Zero"
test$key_cat[test$key==1]<-"One"
test$key_cat[test$key==2]<-"Two"
test$key_cat[test$key==3]<-"Three"
test$key_cat[test$key==4]<-"Four"
test$key_cat[test$key==5]<-"Five"
test$key_cat[test$key==6]<-"Six"
test$key_cat[test$key==7]<-"Seven"
test$key_cat[test$key==8]<-"Eight"
test$key_cat[test$key==9]<-"Nine"
test$key_cat[test$key==10]<-"Ten"
test$key_cat[test$key==11]<-"Eleven"

#Taking the numbers of words in each song
test$song1 <- gsub("\\([^()]*\\)", "", test$song)
test$song_counter <- str_count(test$song1, "\\s+") + 1
#Transforming the count to characters to make them categorical
test$song_count[test$song_counter==1]<-"One"
test$song_count[test$song_counter==2]<-"Two"
test$song_count[test$song_counter==3]<-"Three"
test$song_count[test$song_counter==4]<-"Four"
test$song_count[test$song_counter==5]<-"Five"
test$song_count[test$song_counter==6]<-"Six"
test$song_count[test$song_counter==7]<-"Seven"
test$song_count[test$song_counter >=8]<-"Eight Plus"

#Taking the numbers of words in each performer
test$performer_counter <- str_count(test$performer, "\\s+") + 1
#Transforming the count to characters to make them categorical
test$performer_count[test$performer_counter==1]<-"One"
test$performer_count[test$performer_counter==2]<-"Two"
test$performer_count[test$performer_counter==3]<-"Three"
test$performer_count[test$performer_counter>=4]<-"Four Plus"


#Categorizing the genres
test$pop_genre <- grepl("pop", test$genre, ignore.case = TRUE) & !grepl("hip hop", test$genre, ignore.case = TRUE)
test$rock_genre <- grepl("rock", test$genre, ignore.case = TRUE)
test$adult_genre <- grepl("adult standard", test$genre, ignore.case = TRUE)
test$soul_genre <- grepl("soul", test$genre, ignore.case = TRUE)
test$rap_genre <- grepl("rap", test$genre, ignore.case = TRUE)
test$country_genre <- grepl("country", test$genre, ignore.case = TRUE)
test$hiphop_genre <- grepl("hip hop", test$genre, ignore.case = TRUE)
test$electro_genre <- grepl("electro", test$genre, ignore.case = TRUE)
test$latin_genre <- grepl("latin", test$genre, ignore.case = TRUE)

#Converting booleans to categorical
test$mode<-as.factor(test$mode)

test <- test %>% 
  mutate_if(is.logical, as.factor)
```

For Boosting model
```{r}
dat$performer_counter<-as.factor(dat$performer_counter)
dat$song_counter<-as.factor(dat$song_counter)
dat$key<-as.factor(dat$key)
dat$time_signature<-as.factor(dat$time_signature)

test$performer_counter<-as.factor(test$performer_counter)
test$song_counter<-as.factor(test$song_counter)
test$key<-as.factor(test$key)
test$time_signature<-as.factor(test$time_signature)
```


To check results we will use index
```{r}
set.seed(1010) 
train_index <- sample.split(dat$rating, SplitRatio = 0.8) 
train1 <- dat[train_index,] 
test1 <- dat[!train_index,]
```

Needed names:

track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes + time_sig_cat + key_cat + song_count +  performer_count + pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre

Names needed for training:

id,performer,song,genre,track_duration,track_explicit,danceability,energy,key,loudness,mode,speechiness,acousticness,instrumentalness,liveness,valence,tempo,time_signature,rating,track_minutes,time_sig_cat,key_cat,song1,song_counter,song_count,performer_counter,performer_count,pop_genre,rock_genre,adult_genre,soul


_randomForest_

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

set.seed(1031)
tuned_forest_m <- randomForest(rating~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes + time_sig_cat + key_cat + song_count +  performer_count + pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre,
                      data = train1,
                      mtry = 3, 
                      ntree = 300,
                      allowParallel = TRUE,
                      nodesize = 5,
                      replace = FALSE,
                      splitrule ="variance")

stopCluster(cl)

pred_tuned_forest_m <- predict(tuned_forest_m, newdata = test1)
rmse_train_tuned_forest_m = sqrt(mean((pred_tuned_forest_m - test1$rating)^2)); rmse_train_tuned_forest_m

```

```{r}
submissionFile_10test1 = data.frame (id = test1$id , rating = pred_tuned_forest_m)
View(submissionFile_10test1[1:20,])
```


```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)


set.seed(1031)
forest_ranger = ranger(rating~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes + time_signature + key + song_counter +  performer_counter + pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre,
               data = train1,
               mtry=3,
               min.node.size = 5,
               splitrule= "variance",
               num.trees= 500,
               allowParallel = TRUE,
               replace = FALSE)

stopCluster(cl)

pred_forest_ranger = predict(forest_ranger, data = test1, num.trees = 300)
rmse_forest_ranger = sqrt(mean((pred_forest_ranger$predictions-test1$rating)^2)); rmse_forest_ranger

```

```{r}
submissionFile_10test2 = data.frame (id = test1$id , rating = pred_forest_ranger$prediction)
View(submissionFile_10test2[1:20,])
```


_Boosting model_


```{r}
library(gbm)
set.seed(1031)
boost = gbm(rating~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes  +  pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre+
              performer_counter+song_counter+key+time_signature,
            data=train1,
            distribution="gaussian",
            n.trees = 300,
            interaction.depth = 12,
            shrinkage = 0.051,
            n.minobsinnode=2)

pred_boost = predict(boost, newdata = test1, n.trees = 200)
rmse_boost_train = sqrt(mean((pred_boost-test1$rating)^2)); rmse_boost_train
```

```{r}
submissionFile_10test3 = data.frame (id = test1$id , rating = pred_boost)
View(submissionFile_10test3[1:20,])
```


```{r}
joined_df1 <- left_join(submissionFile_10test2, submissionFile_10test3, by = "id")

submissionFile_ensemble <- joined_df1 %>%
  mutate(rating1 = (rating.x+ rating.y)/2) %>% select(id, rating=rating1)

rmse_ensemble = sqrt(mean((submissionFile_ensemble$rating-test1$rating)^2)); rmse_ensemble

```
_REDO WITH ACTUAL DATA_

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

set.seed(1031)
tuned_forest_m <- randomForest(rating ~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes + time_sig_cat + key_cat + song_count +  performer_count + pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre,
                      data = dat,
                      mtry = 3, 
                      ntree = 300,
                      allowParallel = TRUE,
                      nodesize = 5,
                      replace = FALSE,
                      splitrule ="variance")

stopCluster(cl)
```
```{r}
pred_tuned_forest_m <- predict(tuned_forest_m,  newdata = test )

```

```{r}
submissionFile_10a = data.frame (id = test$id , rating = pred_tuned_forest_m)
View(submissionFile_10a[1:20,])
```


```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)


set.seed(1031)
forest_ranger = ranger(rating~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes + time_signature + key + song_counter +  performer_counter + pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre,
               data = dat,
               mtry=3,
               min.node.size = 5,
               splitrule= "variance",
               num.trees= 500,
               allowParallel = TRUE,
               replace = FALSE)

stopCluster(cl)

```
```{r}
pred_forest_ranger = predict(forest_ranger, data = test, num.trees = 300)
rmse_forest_ranger = sqrt(mean((pred_forest_ranger$predictions-test$rating)^2)); rmse_forest_ranger

```

```{r}
submissionFile_10b = data.frame (id = test$id , rating = pred_forest_ranger$prediction)
View(submissionFile_10b[1:20,])
```


_Boosting model_


```{r}
library(gbm)
set.seed(1031)
boost = gbm(rating~ track_explicit + danceability + energy + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo +  track_minutes  +  pop_genre + rock_genre + adult_genre + soul_genre + rap_genre + country_genre + hiphop_genre + electro_genre + latin_genre+
              performer_counter+song_counter+key+time_signature,
            data=dat,
            distribution="gaussian",
            n.trees = 300,
            interaction.depth = 12,
            shrinkage = 0.051,
            n.minobsinnode=2)

pred_boost = predict(boost, newdata = test, n.trees = 200)
rmse_boost_train = sqrt(mean((pred_boost-test1$rating)^2)); rmse_boost_train
```

```{r}
submissionFile_10c = data.frame (id = test$id , rating = pred_boost)
View(submissionFile_10c[1:20,])
```


```{r}
joined_df1 <- left_join(submissionFile_10a, submissionFile_10c, by = "id")

submissionFile_10ac <- joined_df1 %>%
  mutate(rating1 = (rating.x+ rating.y)/2) %>% select(id, rating=rating1)

```


```{r}
write.csv( submissionFile_10a , 'yosuecanaviri_submission_10a.csv', row.names =F)
write.csv( submissionFile_10b , 'yosuecanaviri_submission_10b.csv', row.names =F)
write.csv( submissionFile_10ac , 'yosuecanaviri_submission_10ac.csv', row.names =F)

```