---
title: "apan5200kaggle1"
author: "Yosue Canaviri"
date: "2023-03-09"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(doParallel)
library(foreach)
```

```{r}
dat <-read.csv("analysisData.csv") # Training set
test <-read.csv("scoringData.csv") # Testing Set
```

Saving clean data:

```{r}
clean_dat <-read.csv("analysisData.csv") # Training set
clean_test <-read.csv("scoringData.csv") # Testing Set

dat<- dat%>%mutate(track_minutes=track_duration/(1000*60))


dat$rating<-pmax(0,dat$rating)


dat$time_sig_cat<- character(length= nrow (dat))
dat$time_sig_cat[dat$time_signature==4]<-"Four"
dat$time_sig_cat[dat$time_signature==3]<-"Three"
dat$time_sig_cat[dat$time_signature %in% c(0,1,5)]<-"Other"
```

  
```{r}
#Mutating data to minutes:
test<- test%>%mutate(track_minutes=track_duration/(1000*60))


test$time_sig_cat<- character(length= nrow (test))
test$time_sig_cat[test$time_signature==4]<-"Four"
test$time_sig_cat[test$time_signature==3]<-"Three"
test$time_sig_cat[test$time_signature %in% c(0,1,5)]<-"Other"
```

Model TUNED

```{r}
set.seed(1031)
tree_8a = rpart( rating~ danceability + energy + loudness + speechiness + acousticness + instrumentalness +liveness+tempo+time_signature+track_explicit+key+mode+valence + track_minutes, data=dat, method = 'anova',
            
               control = rpart.control(cp = 0.0005)
               )
```

```{r}
tree_8b = rpart( rating~ track_minutes + track_explicit + loudness + acousticness + danceability +
                 energy + valence , data=dat, method = 'anova',
                 control = rpart.control(cp = 0.0005)
               )
```

```{r}
pred_8a = predict(tree_8a)
rmse_8a = sqrt(mean((pred_8a - dat$rating)^2)); rmse_8a

pred_8b = predict(tree_8b)
rmse_8b = sqrt(mean((pred_8b - dat$rating)^2)); rmse_8b
```
TRAINING the model with tuned CP

```{r}
#Using this to do a 5 fold validation from 0-0.01 in steps of 0.0001
tuneGrid = expand.grid(cp = seq(0,0.01,0.00001))
trControl = trainControl(method = 'cv',number = 10)
#Need to set.seed
set.seed(1031)
#adding the tuneGrid
tree_cv = train(rating~ danceability + energy + loudness + speechiness + acousticness + instrumentalness +liveness+tempo+time_signature+track_explicit+key+mode+valence + track_minutes, 
                data=dat, 
               method = "rpart",
               trControl = trControl,
               tuneGrid = tuneGrid )
```
```{r}
tree_cv$results
tree_cv$bestTune
```

Tree model with new CP

```{r}
tuned_tree = rpart(rating~ danceability + energy + loudness + speechiness + acousticness + instrumentalness +liveness+tempo+time_signature+track_explicit+key+mode+valence + track_minutes, 
                data=dat ,
               method = "anova",
               cp= 0.00132)
pred_train_tuned = predict(tuned_tree, newdata=test)#model 9_1
rmse_train_tuned = sqrt(mean((pred_train_tuned - dat$rating)^2)); rmse_train_tuned
```

```{r}
submissionFile_9a = data.frame (id = test$id , rating = pred_train_tuned)
View(submissionFile_9a[1:20,])
```

Fitting a Bag model using bagging()

```{r}
library(ipred)
```

```{r}
set.seed(1031) 
bag_ipred = bagging(rating~ danceability + energy + loudness + speechiness + acousticness + instrumentalness +liveness+tempo+time_signature+track_explicit+key+mode+valence + track_minutes, 
                data=dat ,
               nbagg=300 )
#RMSE for tuned model
pred_train_bag_ipred = predict(bag_ipred, newdata=test)#MODEL 9_2
rmse_train_bag_ipred = sqrt(mean((pred_train_bag_ipred - dat$rating)^2)); rmse_train_bag_ipred
```

```{r}
submissionFile_9b = data.frame (id = test$id , rating = pred_train_bag_ipred)
View(submissionFile_9b[1:20,])
```

### Question 5: 
Fitting a Bag model using randomForest
```{r}

```

```{r}
cl <- makeCluster(num_cores)
registerDoParallel(cl)


set.seed(1031) 
bag10 = randomForest(rating~ danceability + energy + loudness + speechiness + acousticness + instrumentalness +liveness+tempo+time_signature+track_explicit+key+mode+valence + track_minutes, 
                data=dat  ,
               mtry = 7,
               ntree=250,
               allowParallel = TRUE,
              nodesize = 10,
               replace = FALSE,
               progress = rfProgress)

stopCluster(cl)

#RMSE for tuned model
pred_train_bag = predict(bag10, newdata=test) #Model 9 _3
rmse_train_bag = sqrt(mean((pred_train_bag - dat$rating)^2)); rmse_train_bag
```

```{r}
submissionFile_9c = data.frame (id = test$id , rating = pred_train_bag)
View(submissionFile_9c[1:20,])
```


## Assignment 8c

### Question 1:

```{r}
cl <- makeCluster(num_cores)
registerDoParallel(cl)

set.seed(1031)
forest <- randomForest(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes, 
                       data = dat,  
                       ntree = 200, 
                       allowParallel = TRUE,
                       nodesize = 10,
                       replace = FALSE,
                       progress = rfProgress)


stopCluster(cl)
```


```{r}
pred_forest = predict(forest, newdata=test)
rmse_forest = sqrt(mean((pred_forest - dat$rating)^2)); rmse_forest
```

```{r}
submissionFile_9d = data.frame (id = test$id , rating = pred_forest)
View(submissionFile_9d[1:20,])
```

### Question 2: 

Tuning number of predictors OR _mtry_
```{r}
library(doParallel)

cl <- makeCluster(detectCores())
registerDoParallel(cl)

ctrl <- trainControl(method = "cv", number = 5, allowParallel = TRUE, verboseIter = TRUE)


grid <- expand.grid(mtry = 1:14)

set.seed(1031)
rf_model <- train(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
                  data = dat,
                  method = "rf",
                  ntree = 250,
                  trControl = ctrl,
                  tuneGrid = grid)

stopCluster(cl)
```

```{r}
forest_mtry = rf_model
```


```{r}
forest_mtry
forest_mtry$bestTune$mtry
```
_Model 9e_

```{r}

library(doParallel)

cl <- makeCluster(detectCores())
registerDoParallel(cl)

set.seed(1031)
tuned_forest <- randomForest(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
                      data = dat,
                      mtry = 2, 
                      ntree = 300,
                      allowParallel = TRUE,
                      nodesize = 25,
                      replace = FALSE,
                      splitrule ="variance")


stopCluster(cl)

pred_tuned_forest <- predict(tuned_forest, newdata = test)
rmse_train_tuned_forest = sqrt(mean((pred_tuned_forest - dat$rating)^2)); rmse_train_tuned_forest

```

```{r}
submissionFile_9e = data.frame (id = test$id , rating = pred_tuned_forest)
View(submissionFile_9e[1:20,])
```

Trying with splitrule _extratrees_

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

set.seed(1031)
tuned_forest_e <- randomForest(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
                      data = dat,
                      mtry = 3, 
                      ntree = 300,
                      allowParallel = TRUE,
                      nodesize = 20,
                      replace = FALSE,
                      splitrule ="extratrees")

stopCluster(cl)

pred_tuned_forest_e <- predict(tuned_forest_e, newdata = test)
rmse_train_tuned_forest_e = sqrt(mean((pred_tuned_forest_e - dat$rating)^2)); rmse_train_tuned_forest_e

```

```{r}
submissionFile_9f = data.frame (id = test$id , rating = pred_tuned_forest_e)
View(submissionFile_9f[1:20,])
```

Trying with splitrule _extratrees_

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

set.seed(1031)
tuned_forest_m <- randomForest(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
                      data = dat,
                      mtry = 2, 
                      ntree = 200,
                      allowParallel = TRUE,
                      nodesize = 5,
                      replace = FALSE,
                      splitrule ="maxstat")


stopCluster(cl)

pred_tuned_forest_m <- predict(tuned_forest_m, newdata = test)
rmse_train_tuned_forest_m = sqrt(mean((pred_tuned_forest_m - dat$rating)^2)); rmse_train_tuned_forest_m

```

```{r}
submissionFile_9g = data.frame (id = test$id , rating = pred_tuned_forest_m)
View(submissionFile_9g[1:20,])
```



Using _ranger()_

```{r}
library(ranger)
```
```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)


set.seed(1031)
forest_ranger = ranger(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
               data = dat,
               mtry=2,
               min.node.size = 3,
               splitrule= "variance",
               num.trees= 300,
               allowParallel = TRUE,
               replace = FALSE)

stopCluster(cl)

pred_forest_ranger = predict(forest_ranger, data = test, num.trees = 300)

```

```{r}
submissionFile_9h = data.frame (id = test$id , rating = pred_forest_ranger$prediction)
View(submissionFile_9h[1:20,])
```

_TRAINING RANGER_

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

trControl3=trainControl(method="cv",number=5)
tuneGrid3 = expand.grid(mtry=1:14, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(1,2,3,4,5,10,20,50,100))
set.seed(1031)
cv_forest_ranger = train(rating ~ danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + track_explicit + key + mode + valence + track_minutes,
                data = dat,
                method="ranger",
                num.trees=250,
                parallel="multicore",
                trControl=trControl3,
                tuneGrid=tuneGrid3 )

cv_forest_ranger$bestTune
```

```{r}
set.seed(1031)
tuned_forest_ranger = ranger(rating ~ company_location+country_of_bean_origin+cocoa_percent+ingredient_Beans+ingredient_Sugar+
               ingredient_Cocoa_Butter+ingredient_vanilla+ingredient_lecithin+ingredient_salt+sweet+nutty+cocoa+
               roasty+earthy+creamy+sandy+fatty+floral+intense+spicy+sour+vanilla+fruit+molasses+woody+sticky+coffee+
               rich+dried.fruit,
                          data=train,
                          num.trees = 200, 
                          mtry=cv_forest_ranger$bestTune$mtry, 
                          min.node.size = cv_forest_ranger$bestTune$min.node.size, 
                          splitrule = cv_forest_ranger$bestTune$splitrule)

pred_tuned_forest_ranger = predict(tuned_forest_ranger, data =train, num.trees = 200)
rmse_tuned_forest_ranger = sqrt(mean((pred_tuned_forest_ranger$predictions-train$rating)^2)); rmse_tuned_forest_ranger
```

### Question 6:
_Boosting model_

```{r}
library(gbm)
set.seed(1031)
boost = gbm(rating~danceability + energy + loudness + speechiness + acousticness + instrumentalness + liveness + tempo + time_signature + key + mode + valence + track_minutes,
            data=dat,
            distribution="gaussian",
            n.trees = 200,
            interaction.depth = 3,
            shrinkage = 0.081,
            n.minobsinnode=2)

pred_boost = predict(boost, newdata = test, n.trees = 200)
rmse_boost_train = sqrt(mean((pred_boost-dat$rating)^2)); rmse_boost_train
```

```{r}
submissionFile_9i = data.frame (id = test$id , rating = pred_boost)
View(submissionFile_9i[1:20,])
```


```{r}
joined_df1 <- left_join(submissionFile_9i, submissionFile_9h, by = "id")


joined_df2 <- left_join(joined_df1, submissionFile_9f, by = "id")


final_df <- joined_df2

submissionFile_ensemble <- final_df %>%
  mutate(rating1 = (rating.x + rating.y + rating) / 3) %>% select(id, rating=rating1)

```

_Submission for Model 9_
```{r}
submissionFile9 = data.frame (id = test$id , rating = pred8)
write.csv( submissionFile8 , 'yosuecanaviri_submission8.csv', row.names =F)
```

```{r}
write.csv( submissionFile_9a , 'yosuecanaviri_submission_9a.csv', row.names =F)
write.csv( submissionFile_9b , 'yosuecanaviri_submission_9b.csv', row.names =F)
```
```{r}
write.csv( submissionFile_9c , 'yosuecanaviri_submission_9c.csv', row.names =F)
write.csv( submissionFile_9d , 'yosuecanaviri_submission_9d.csv', row.names =F)
write.csv( submissionFile_9f , 'yosuecanaviri_submission_9f.csv', row.names =F)
write.csv( submissionFile_9g , 'yosuecanaviri_submission_9g.csv', row.names =F)

write.csv( submissionFile_9h , 'yosuecanaviri_submission_9h.csv', row.names =F)
write.csv( submissionFile_9i , 'yosuecanaviri_submission_9i.csv', row.names =F)

write.csv( submissionFile_ensemble , 'yosuecanaviri_submission_ensemble.csv', row.names =F)
```
